Peridynamic Beams: \\ A Non-ordinary, State Based Model

### 2014-02-17 BeamPaper.tex

\section{Introduction}
The final goal of many mechanical engineering analyses is the prediction and description of material failure.
When processes such as fracture are modeled, the differential equations of classical mechanics are ill-defined at the resulting discontinuities.
A peridynamic formulation of continuum mechanics casts material behavior in terms of integral functions of displacement, so that discontinuities evolve naturally and require no special treatment.
Various peridynamic models capture the material behavior of 3-dimensional solid objects, but would be very expensive to implement for a thin plate or beam.
Other peridynamic models capture tension and compression in 1D bars and 2D plates, but these models do not resist transverse displacement.
A recent paper by Taylor and Steigmann \cite{taylor2013two} reduces a bond based 3D plate to two dimensions with an integral through the plate's thickness.

This paper begins to remedy that lack in the simplest case by presenting a peridynamic equivalent to an Euler-Bernoulli beam.
In addition to directly modeling a thin beam in bending, the simple beam case lays the theoretical framework for more complex peridynamic beam, plate, and shell bending models.
Because many analyses of interest are partly or wholly comprised of these types of features, their development is an important addition to the capabilities of peridynamic analysis.
%
%The second section of this paper provides a brief introduction to peridynamics, including state based models.
%The third section presents the state based beam model and demonstrates equivalence to classical Euler-Bernoulli beam theory in the limit of shrinking nonlocality.
%The fourth section demonstrates the model's equivalence to Eringen's nonlocal elasticity for small peridynamic horizons, and compares other gradient models.
%The fifth section demonstrates the beam model with simple test cases.
The remainder of this introduction reviews other nonlocal work and provides a brief introduction to peridynamics, including state based models.
\Cref{sec:ModelDev} presents the state based beam model and demonstrates equivalence to classical Euler-Bernoulli beam theory in the limit of shrinking nonlocality.
\Cref{sec:EringenCompare} demonstrates the model's equivalence to Eringen's nonlocal elasticity for small peridynamic horizons.
\Cref{sec:Numerical} demonstrates the beam model with simple test cases.
%
\subsection{Nonlocal Beam Models}
\label{sec:NLbeams}
%
Nonlocal elasticity generally allows for forces at a point that are dependent on the material configuration of an entire body, rather than the configuration at that point.\cite{eringen1972nonlocal}
While long-range forces are obvious at the molecular model, material at larger scales is conventionally modeled as though internal forces are local or contact forces.\cite{kroner1967elasticity}
The result of such approximation is accurate for deformations that are homogenous, but introduces some inaccuracy for inhomogenous deformations like the propagation of waves with short wavelengths.
One way to distinguish between homogenous and inhomogenous deformations is to incorporate higher-order gradients of deformation.
While stress in classical elasticity is a function of the (first) gradient of deformation, Eringen's formulation of a nonlocal modulus in \cite{eringen1983differential} approximates a weighted sum of the first and second order gradients.
This introduces a length scale to the model and has the effect of smearing out local deformation inhomogeneities over the surrounding material, while maintaining the conventional result for homogenous deformations.

@@

\subsection{Weighting Function and Perfect Plasticity}
\label{sec:WeightFunction}
The weighting function \(\omega(\xi)\) describes the relative contribution of each bond-pair, and can be defined according to physical or mathematical considerations. 
Consider an Euler beam in bending with curvature \(\kappa\). 
Fibers running parallel to the neutral axis of the beam are stretched in proportion to their distance from the neutral axis, with strain \(\epsilon = y\kappa\). 
If the fibers are linearly elastic, then the axial stress at each location is \(\sigma = E\epsilon = Ey\kappa\), and the contribution to supported moment \(dM = \kappa E y^2 dA\). 
By comparing the formulations for the moments carried by the Euler beam in \cref{fig:EulerBending} and those of the bond-pair beam in \cref{fig:BPBending}, we see some definite parallels.
%
\begin{figure}[h]
\subinputfrom{\diagrampath}{EulerBending.eps_tex}
\caption{Euler beam moment contribution}
\label{fig:EulerBending}
\end{figure}
%%
\begin{figure}[h]
\subinputfrom{\diagrampath}{BondPairBending_edit.eps_tex}
\caption{Bond-pair moment contribution}
\label{fig:BPBending}
\end{figure}
%
\begin{align}
M_\text{classical}&=\int_{-\frac{t}{2}}^{\frac{t}{2}} \sigma \; y \; dA && = \int_{-\frac{t}{2}}^{\frac{t}{2}} E \kappa \; y^2 \; b(y) dy\notag \\
%
M_\text{PD}&=\int_{-\delta}^{\delta} \vstate{T}{}{\xi}\; \xi \; d\xi & = \int_{-\delta}^{\delta} \alpha \frac{\sin(\Delta\theta)}{|\xi|} \; \xi \; \omega(\xi) d\xi\: & \approx \int_{-\delta}^{\delta} \alpha \kappa |\xi| \; \omega(\xi) d\xi\notag
\end{align}
%
The term \(y\) is the distance from the beam's neutral axis and \(b(y)\) is the width of the beam at that distance from the neutral axis. 
The similarity between classical and peridynamic moment formulations suggests a possible formulation for the weighting function:
%
\begin{equation}
\omega(\xi) = |\xi| b\left(y\right) \text{at}\; y=\frac{\xi}{\delta} \frac{t}{2}\notag
\end{equation}
%
\begin{figure}[h]
\minipage[t]{0.5\textwidth}
 \centering
  \subinputfrom{\diagrampath}{WeightProfile_Uniform.eps_tex}
\caption{Weight function for a beam of rectangular cross-section}
\label{fig:WeightProfileUniform}
\endminipage\hfill
\minipage[t]{0.5\textwidth}%
  \centering
  \subinputfrom{\diagrampath}{WeightProfile_Ibeam.eps_tex}
\caption{Weight function for an I-beam}
\label{fig:WeightProfileIbeam}
\endminipage
\end{figure}
%
This weight function analogizes the relative contributions of bond pairs of different lengths to the relative contributions of fibers at different distances from the centerline. 
Examples are shown in \cref{fig:WeightProfileUniform,fig:WeightProfileIbeam}.
While this weighting function offers no advantages over a uniform weight function in the case of the linearly elastic beam, it offers a way to model advancing plasticity.

>>>>

\section{Introduction}
A goal of many mechanical engineering analyses is the prediction and description of material failure.
When processes such as fracture are modeled, the partial-differential equations of classical mechanics are ill-defined at the resulting discontinuities in displacement.  A peridynamic formulation of continuum mechanics casts material behavior in terms of integral functions of displacement (as opposed to gradients of displacement), so that discontinuities evolve naturally and require no special treatment.  Various peridynamic material models capture the material behavior of 3-dimensional solid objects \cite{silling2007peridynamic, silling2005meshfree, gerstle2007peridynamic}, but would be very expensive to implement for a thin plate or beam, as the thru-thickness discretization requirement to properly capture resistance to bending would cause an inordinate amount of computational nodes along the length in a long, slender structural object.  Other peridynamic models capture tension and compression in 1D bars and 2D plates, but these models do not resist transverse displacement {\color{red} [Ref]}.  A recent paper by Taylor and Steigmann \cite{taylor2013two} reduces a bond based 3D plate to two dimensions with an integral through the plate's thickness, to create a model that can represent thin structures, but fails to properly {\color{red} XXX}.

This paper presents a peridynamic equivalent to an Euler-Bernoulli beam, along with a methodology for representing non-uniform cross-sections, plastic behavior, and failure.  In addition to directly modeling a thin beam in bending, the simple beam case lays the theoretical framework for more complex peridynamic beam, plate, and shell bending models.
Because many analyses of interest are partly or wholly comprised of these types of features, their development is an important addition to the capabilities of peridynamic analysis.
%
%The second section of this paper provides a brief introduction to peridynamics, including state based models.
%The third section presents the state based beam model and demonstrates equivalence to classical Euler-Bernoulli beam theory in the limit of vanishing nonlocality.
%The fourth section demonstrates the model's equivalence to Eringen's nonlocal elasticity for small peridynamic horizons, and compares other gradient models.
%The fifth section demonstrates the beam model with simple test cases.
The remainder of this introduction reviews other nonlocal work and provides a brief introduction to peridynamics, including state based models.
\Cref{sec:ModelDev} presents the state based beam model and demonstrates equivalence to classical Euler-Bernoulli beam theory in the limit of shrinking nonlocality.
\Cref{sec:EringenCompare} demonstrates the model's equivalence to Eringen's nonlocal elasticity for small peridynamic horizons.
\Cref{sec:Numerical} demonstrates the beam model with simple test cases.
%
\subsection{Nonlocal Beam Models}
\label{sec:NLbeams}
%
Nonlocal elasticity generally allows for forces at a point that are dependent on the material configuration of an entire body, rather than the configuration at that point \cite{eringen1972nonlocal}.  While long-range forces are obvious at the molecular model, material at larger scales is conventionally modeled as though internal forces are local or contact forces \cite{kroner1967elasticity}.
The result of such approximation is accurate for deformations that are homogenous, but introduces some inaccuracy for inhomogenous deformations like the propagation of waves with short wavelengths.
One way to distinguish between homogenous and inhomogenous deformations is to incorporate higher-order gradients of deformation.
While stress in classical elasticity is a function of the (first) gradient of deformation, Eringen's formulation of a nonlocal modulus in \cite{eringen1983differential} approximates a weighted sum of the first and second order gradients.
This introduces a length scale to the model and has the effect of smearing out local deformation inhomogeneities over the surrounding material, while maintaining the conventional result for homogenous deformations.

@@

\subsection{Weighting Function and Perfect Plasticity}
\label{sec:WeightFunction}
The weighting function \(\omega(\xi)\) describes the relative contribution of each bond-pair, and can be defined according to physical or mathematical considerations. 
Consider a classical Euler-Bernoulli beam in bending with curvature \(\kappa\). 
Fibers running parallel to the neutral axis of the beam are stretched in proportion to their distance from the neutral axis, with strain \(\epsilon = y\kappa\). 
If the fibers are linearly elastic, then the axial stress at each location is \(\sigma = E\epsilon = Ey\kappa\), and the contribution to supported moment \(dM = \kappa E y^2 dA\). 
By comparing the formulations for the moments carried by the Euler beam in \cref{fig:EulerBending} and those of the bond-pair beam in \cref{fig:BPBending}, we see some definite parallels.
%
\begin{figure}[h]
\subinputfrom{\diagrampath}{EulerBending.eps_tex}
\caption{Euler beam moment contribution}
\label{fig:EulerBending}
\end{figure}
%%
\begin{figure}[h]
\subinputfrom{\diagrampath}{BondPairBending_edit.eps_tex}
\caption{Bond-pair moment contribution}
\label{fig:BPBending}
\end{figure}
%
\begin{align}
M_\text{classical}&=\int_{-\frac{t}{2}}^{\frac{t}{2}} \sigma \; y \; dA && = \int_{-\frac{t}{2}}^{\frac{t}{2}} E \kappa \; y^2 \; b(y) dy\notag \\
%
M_\text{PD}&=\int_{-\delta}^{\delta} \vstate{T}{}{\xi}\; \xi \; d\xi & = \int_{-\delta}^{\delta} \alpha \frac{\sin(\Delta\theta)}{|\xi|} \; \xi \; \omega(\xi) d\xi\: & \approx \int_{-\delta}^{\delta} \alpha \kappa |\xi| \; \omega(\xi) d\xi\notag
\end{align}
%
The term \(y\) is the distance from the beam's neutral axis and \(b(y)\) is the width of the beam at that distance from the neutral axis. 
The similarity between classical and peridynamic moment formulations suggests a possible formulation for the weighting function:
%
\begin{equation}
\omega(\xi) = |\xi| b\left(y\right) \text{at}\; y=\frac{\xi}{\delta} \frac{t}{2}\notag
\end{equation}
%
\begin{figure}[h]
\minipage[t]{0.5\textwidth}
 \centering
  \subinputfrom{\diagrampath}{WeightProfile_Uniform.eps_tex}
\caption{Weight function for a beam of rectangular cross-section}
\label{fig:WeightProfileUniform}
\endminipage\hfill
\minipage[t]{0.5\textwidth}%
  \centering
  \subinputfrom{\diagrampath}{WeightProfile_Ibeam.eps_tex}
\caption{Weight function for an I-beam}
\label{fig:WeightProfileIbeam}
\endminipage
\end{figure}
%
This weight function analogizes the relative contributions of bond pairs of different lengths to the relative contributions of fibers at different distances from the centerline. 
Examples are shown in \cref{fig:WeightProfileUniform,fig:WeightProfileIbeam}.
    While this weighting function offers no advantages over a uniform weight function in the case of the linearly elastic beam, it offers a way to model advancing plasticity.
